---
title: "viz_part_01"
output: github_document
---


```{r, echo = FALSE,message=FALSE}
library(tidyverse)
library(ggridges)
```

import weather dataset

```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USC00519397", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USC00519397 = "Waikiki_HA",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
```

## basic scatterplot

We’ll use the weather_df data throughout, so we’ll move straight into defining aesthetic mappings. To create a basic scatterplot, we need to map variables to the X and Y coordinate aesthetics:

```{r}
ggplot(weather_df, aes(x = tmin, y = tmax))+
  geom_point()

```

let's make the same scatterplot, but different

```{r}
weather_df %>% 
  drop_na() %>% 
  filter(name=="CentralPark_NY") %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point()
```

let's keep making the same plot but different

```{r}
weather_scatterplot = 
  weather_df %>% 
  drop_na() %>% 
  ggplot(aes(x=tmin,y=tmax))

weather_scatterplot + 
  geom_point()
```

##let's fancy this up

```{r}

weather_df %>% 
  ggplot(aes(x=tmin,y=tmax,color=name)) +
  geom_point()

weather_df %>% 
  ggplot(aes(x=tmin,y=tmax)) +
  geom_point(aes(color=name))+
  geom_smooth()

weather_df %>% 
  ggplot(aes(x=tmin,y=tmax,color=name)) +
  geom_point(alpha=.3)+ #alpha transparancy 0~1
  geom_smooth(se=FALSE)


```

maybe make seperate panels

```{r}
weather_df %>% 
  ggplot(aes(x=tmin,y=tmax,color=name)) +
  geom_point(alpha=.3)+ #alpha transparancy 0~1
  geom_smooth(se=FALSE)+
  facet_grid(. ~ name) #row ~ column

weather_df %>% 
  ggplot(aes(x=tmin,y=tmax,color=name)) +
  geom_point(alpha=.3)+ #alpha transparancy 0~1
  geom_smooth(se=FALSE)+
  facet_grid(name ~ .) #row ~ column, . is a placeholder
```

`tmax` vs `tmin` is boring. spicing it up

```{r}
weather_df %>% 
  ggplot(aes(x=date,y=tmax,color=name))+
  geom_point(aes(size=prcp),alpha=.3)+
  geom_smooth(se=FALSE)+
  facet_grid(.~name)+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))#rotate x axis labels 
```

## some quick stuff
```{r}
weather_df %>% 
  ggplot(aes(x=tmin,y=tmax))+
  geom_hex()
```

## univariate plots

histograms, barplots, boxplots, violins

```{r}
weather_df %>% 
  ggplot(aes(x=tmax,fill=name))+
  geom_histogram()+
  facet_grid(.~name)
```

more options

```{r}
weather_df %>% 
  ggplot(aes(x=tmax,fill=name,color=name)) +
  geom_density(alpha=.1)
```

boxplots

```{r}
weather_df %>% 
  ggplot(aes(x=name,y=tmax,fill=name))+
  geom_boxplot()
```

violin

```{r}
weather_df %>% 
  ggplot(aes(x=name,y=tmax,fill=name))+
  geom_violin(alpha=.3)
```

OR

ridges plot for more groups comparison, same as violin 

```{r}
weather_df %>% 
  ggplot(aes(x=tmax,y=name))+
  geom_density_ridges()
```

## saving and embedding plots

first -- let's save a plot

```{r}
weather_scatterplot = 
  weather_df %>% 
  ggplot(aes(x=tmin,y=tmax,color=name))+
  geom_point(aes(size=prcp),alpha=.3)+
  geom_smooth(se=FALSE)+
  facet_grid(.~name)

weather_scatterplot

ggsave(file="results/weather_scatterplot.pdf", plot=weather_scatterplot, width = 8, height = 5)

```

```{r, fig.width=10}
weather_scatterplot

```



